Habiba

### 2016-05-30 Tanzania/dse.tz.Rmd
Add DSensemble results

# R Markdown

@@

The code below shows how the functions are used and applied to the data for the daily maximum $T_x$. The first chunck of code carries out some preliminary analysis, the second shows how the results are derived, and the third show how the plots are made.

@@

param <- 'tmax'
FUN <- 'mean'
reanalysis <- 'air.mon.mean.nc'
#reanalysis <- 'ERAINT_t2m_mon.nc'
FUNX <- 'mean'
data.path <- 'dse.Tz.NCEP/'

@@

The rank test (Wilcox) indicates reasonable scores, although the rank distribution is not entirely flat. The diagnostic (target plot) for the diagnosis suggest that the trend in the ensemble is higher than in observation (horizontal axis) and that the inter-annual variance (vertical axis) is smaller in the ESD results.

@@

A weaker warming (~2C) is projected in DJF, also with strongest response in the north.  But the rank scores are weak.

@@

The assessment suggest the JJA season is better downscaled, with reasonable rank scores. The projected warming is more uniform over Tanzania, albeit with some local variations.

@@

```{r,message=FALSE,warning=FALSE}
load(paste(data.path,'dse.',param,'.tz.rcp85.son.rda',sep=''))
z <- as.station(dse.t2m.tz.rcp85)
gridmap(z)
figlab('SON')
#dev.copy2pdf(file=paste('map.',param,'.son.pdf',sep=''))
validate(z)
figlab('SON')
#dev.copy2pdf(file=paste('map.',param,'.son.validate.pdf',sep=''))
diagnose(z)
figlab('SON')
#dev.copy2pdf(file=paste('map.',param,'.son.diagnose.pdf',sep=''))

The SON season is associated with strongest warming in the north, and displays a clear north-south gradient. The warming is projected to be 0.7--1.3C by 2099, and the rank scores are moderately good - the distribution of the ranks had a U-shape rather than being flat.

##Check a random station: is is well-represented by the PCA?

@@

The single station data could be well-recovered from the PCA product.

#Implementation - Minimum temperature

@@

Apply the ESD models to implement thedownscaling

@@

The PCA suggest that most of the variance in the station data for daily maximum temperature $T_n$ can be represented by a small number of leading modes. For the SON season, however, the second mode is more prominent, suggesting more complicated structure where spatial variability has less coherency between the stations.

The CCA suggests dissimilar spatial temperature pattern with the higest temporal correlation in the predictand and predictor.

@@

The plume plots look good for Arusha and Musoma, and strongest warming (3.5-4C) is projected for the eastern part of Tanzania. The evaluation of the quality gave moderate results - some discrepancies in terms of the rank scores.

@@

Strongest warming projected for the southeast part of the country (~3C), and the evaluation suggest moderate scores. Again, still some discrepancies.

@@

Most pronounced warming is projected for the souther part (~3C), and the assessment indicates moderate quality with some discrepancies. The ensemble results tend to account for lower inter-annual variability than in the observations and weaker trend (except for one station with higher trend).

@@

```{r,message=FALSE,warning=FALSE}
#load(paste(data.path,'dse.',param,'.tz.rcp85.son.rda',sep=''))
load(paste(data.path,'dse.',param,'.tz.rcp45.son.rda',sep=''))
z <- as.station(dse.t2m.tz.rcp85)
gridmap(z)
figlab('SON')
#dev.copy2pdf(file=paste('map.',param,'.son.pdf',sep=''))
validate(z)
figlab('SON')
#dev.copy2pdf(file=paste('map.',param,'.son.validate.pdf',sep=''))
diagnose(z,new=FALSE)
figlab('SON')
#dev.copy2pdf(file=paste('map.',param,'.son.diagnose.pdf',sep=''))

@@

Strongest projected warming in the south (~3C) and moderate quality.

## Check a random station: is is well-represented by the PCA?

@@

Single station records can be recovered from the PCA.

@@

The NCEP/NCAR reanalysis is probably better than ERAINT for ESD model calibration.

>>>>

## R Markdown

@@

The code below shows how the functions are used and applied to the data for the daily maximum temperature $T_x$. The first chunck of code carries out some preliminary analysis, the second shows how the results are derived, and the third show how the plots are made.

@@

param <- 'tmax'
FUN <- 'mean'
#reanalysis <- 'air.mon.mean.nc'
reanalysis <- 'ERAINT_t2m_mon.nc'
FUNX <- 'mean'
data.path <- 'dse.Tz.ERAINT/'

@@

The rank test (Wilcox) indicates reasonable scores, although the rank distribution is not entirely flat. The diagnostic (target plot) for the diagnosis suggest a reasonable reproduction of trend (horizontal axis) and variance (vertical axis), although four stations were associated with higher variance than suggested by the ensemble and the observed trends were weaker than suggested by the ensemble.

@@

The projected warming was again hiher in the north (~3C), and the evaluation suggested reasonable skill, although not perfect.

@@

In June-August, the projected warming in $T_x$ was found to be greatest the west (~3C), but the assessment of the rank score suggested a poor match between observations and the downscaled ensemble.

@@

```{r,message=FALSE,warning=FALSE}
load(paste(data.path,'dse.',param,'.tz.rcp85.son.rda',sep=''))
z <- as.station(dse.t2m.tz.rcp85)
gridmap(z)
figlab('SON')
#dev.copy2pdf(file=paste('map.',param,'.son.pdf',sep=''))
validate(z)
figlab('SON')
#dev.copy2pdf(file=paste('map.',param,'.son.validate.pdf',sep=''))
diagnose(z)
figlab('SON')
#dev.copy2pdf(file=paste('map.',param,'.son.diagnose.pdf',sep=''))
```

The results suggest strongest warming in the north, but the map is unrealistic with only a latitude-based gradient. The assessment of the ranked score also suggests a weak performance.

## Test of PCA

Check a random station: is it well-represented by the PCA?

@@

The data for Arusha was recovered from the PCA products.

#Minimum temperature

@@

The PCA suggests that most of the variance in $T_n$ was captured by a small number of leading modes. The CCA found that the patterns in the predictand and predictor with highest correlation were not so similar in terms of spatial structure.

The DS applied to the reanalysis found highest scores for MAM, and again, the spatial patterns were dissimilar.

## Apply the ESD models to implement the downscaling

@@

@@

The plume plots for Arusha and Musoma look ok. Strongest warming (~3C) is projected for the eastern part of the country. THe assessment of the downscaled results in terms of ranked score suggest that the observations have a different character to the downscaled results (cold bias).

@@

The projections suggest strongest warming in the southeast (~2C), and again, the evaluation indicates a mismatch between observations and downscaled results. Poor performance.

@@

Strongest warming in the southwest (~3C), and higher ranked scores than the two previous seasons, but the observations have a higher variance than accounted for by the downscaled ensemble and the the trends in the downscaled results are weaker than those observed.

@@

```{r,message=FALSE,warning=FALSE}
load(paste(data.path,'dse.',param,'.tz.rcp85.son.rda',sep=''))
z <- as.station(dse.t2m.tz.rcp85)
gridmap(z)
figlab('SON')
#dev.copy2pdf(file=paste('map.',param,'.son.pdf',sep=''))
validate(z)
figlab('SON')
#dev.copy2pdf(file=paste('map.',param,'.son.validate.pdf',sep=''))
diagnose(z,new=FALSE)
figlab('SON')
#dev.copy2pdf(file=paste('map.',param,'.son.diagnose.pdf',sep=''))

@@

Greatest warming in the southwest, but these results look suspect. The plume-plot for Arusha exhibits a spread that increases over time that spans from cooling to warming.

##Check a random station: is it well-represented by the PCA?

@@

The original data is recovered from the PCA product.

@@

The ERAINT is probably too short for proper model calibration.

### 2016-05-30 Tanzania/tmintest.Rmd
Add DSensemble results

We can interpolate temperatures for the same coordinate as for Mwanza, but when we do the comparison, we need to use standardised values as the predictor represents the mean of daily mean temperatures $\overline{T}$ whereas the station is the mean daily minimum $T[min]$:

>>>>

We can interpolate temperatures for the same coordinate as for Mwanza, but when we do the comparison, we need to use standardised values as the predictor represents the mean of daily mean temperatures $\bar{T}$ whereas the station is the mean daily minimum $T_{min}$:
